Timely Object Detection

### 2012-02-22 writeups/eccv12/eccv12.tex
writeup

% Definitions
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\def\subsectionautorefname{section}
\definecolor{light-gray}{gray}{0.3}
\newcommand{\aside}[1]{\textcolor{light-gray}{\emph{#1}}}
\newcommand{\todo}[1]{\textcolor{red}{\emph{#1}}}
\newcommand{\comment}[1]{}

@@

We deal with a dataset of images $\mathcal{D}$, where each image $\mathcal{I}$ contains at least one, and often multiple, objects.
Each object is labeled with exactly one category label $1$ through $K$.

The multi-class, multi-label \emph{classification} problem asks whether $\mathcal{I}$ contains at least one object of class $i$, for each class $i \in \{1,\dots,K\}$.
The answer for a single label is given with a real-valued confidence by a function $\emph{classify}(\mathcal{I},i)$.
We write the ground truth for an image as $\mathbf{C}=\{C_1,\dots,C_K\}$, where $C_i \in \mathbb{B} = \{0,1\}$ is set to $1$ if an object of class $i$ is present.

@@

We can make the classification problem more difficult by posing the \emph{counting} problem, which asks how many objects of class $i$ are present in $\mathcal{I}$, for each $i$.
This setting is not commonly evaluated; we mention it for its usefulness in later exposition.

Finally, the \emph{detection} problem is to output a list of bounding boxes (sub-images defined by four coordinates), each with a real-valued confidence that it encloses a single instance of an object of class $i$.
The answer for a single class $i$ is given by an algorithm $\emph{detect}(\mathcal{I},i)$, which outputs a list of bounding boxes $B$ and associated confidences.
To highlight the hierarchical structure of these problems, we note that the confidences may be given by $\operatorname{classify}(b,i)$ for each sub-image $b \in B$, and that the correct answer to the detection problem also answers the counting and classification problems.

Here we describe the classification problem setup, although we evaluate performance on both classification and detection tasks.

@@

\begin{table}
\centering
\caption{Summary of the notation used.}
\label{tab:notation}
\begin{tabular}{|l|l|}
	\hline
	$C_i$         & presence of class $i$ \\ 
	$K$           & number of object classes \\ 
	$t$           & time into episode \\ 
	$T_s$, $T_d$  & start and deadline times \\ 
	$\mathcal{A}$ & set of actions \\ 
	$b_j$        	& belief state at step $j$ \\ 
	$\pi$         & policy, $\pi(b) \mapsto a \in \mathcal{A}$ \\ 
	\hline
\end{tabular}\end{table}

@@

Our goal is a multi-class classification policy $\pi$ that takes an image $\mathcal{I}$ and outputs $\{\emph{classify}(1), \dots, \emph{classify}(K)\}$.
The policy repeatedly selects an action $a$ from a set of actions $\mathcal{A}$, executes it, potentially receives a set of observations, and selects the next action.
A dynamic, or ``closed-loop,'' policy bases action selection on observations received from previous actions, exploiting the signal in inter-object and scene context for a maximally efficient path through the classifiers.
This is our goal, and what sets our formulation apart from multi-class systems that evaluate in a fixed order, such as simple cascades \cite{Viola2001} or \todo{add another one}.

The set of actions $\mathcal{A}$ can include per-class classifiers, ``scene context'' functions, and feature computations.
For simplicity of exposition, let $\mathcal{A}$ consist of $K+1$ feature computations $F$ and $2K$ per-class classifiers $L$:
\begin{itemize}
	\item $L^\text{co}_i$ are one-vs-all SVMs on a complex featurization of the image (computed by $F^\text{co}_i$);
	\item $L^\text{sc}_i$ are SVMs on a simple scene-level feature (computed by $F^\text{sc}$).
\end{itemize}

\todo{
Revise the above to make a single action compute all the scene classifiers at once.
}

@@

The system is given two times: the setup time $T_s$ and deadline $T_d$.
Between these times, the system must meet the goal of \emph{Anytime} performance and give the best possible answer to the classification problem at any point.

@@

To meet this goal, the evaluation metric of a policy is defined as the area under the performance vs. time curve between points $T_s$ and $T_d$ (see Figure~\ref{fig:evaluation}).
We evaluate policies by this more robust metric and not simply by the final performance at deadline time for the same reason---robustness---that Average Precision is used instead of a fixed Precision vs. Recall point.

We define the \emph{belief state} $b$ of the decision process by the the distribution over class presence variables $P(\mathbf{C}) = P(C_1, \dots, C_K)$ (writing $P(C_i)$ to mean $P(C_i=1)$), the time into episode $t$, the set of executed actions $\mathcal{O}$, and other variables that will be discussed in \autoref{sec:value}.
\comment{The terminology of ``belief'' comes from reinforcement learning and refers to the distribution $P(\mathbf{C})$.}

An action $a$ can result in a set of observations $o$ that are used to update the belief state.
At the very least, the state records the fact that $a$ has been taken by adding it to the initially empty set $\mathcal{O}$.

A classification \emph{episode} takes an image $\mathcal{I}$ and proceeds from the initial belief state $b_0$ and action $a_0$ to the next pair $(b_1,a_1)$, and so on until $t$ exceeds $T_d$.
At that point, the policy is terminated and a new episode begins on a new image.

The policy's performance is determined by treating the final values of $P(\mathbf{C})$ as the set of confidence scores $\{\emph{classify}(1), \dots, \emph{classify}(K)\}$ for the multi-label classification evaluation (see \autoref{sec:problem_setting}).

\subsection{The value function} \label{sec:value}
Optimal performance results from becoming maximally certain of the correct values $C_i$ as quickly as possible (given the setup time).
As our goal is to pick actions, we want to formulate a function that assigns a value to a potential action, given the current state of the decision process.
We will then be able to define our policy as simply
\begin{equation}
\pi(b) = \argmax_{a \in \mathcal{A} \setminus \mathcal{O}} V(b,a)
\end{equation}

We consider two formulations of the value function: one based on information gain, and one on the classification loss.

\subsubsection{Information gain}
negative residual entropy, following \cite{Gao2011}.

\todo{Write this out, and mention the greedy submodularity argument, to be made in a later section.}

\subsubsection{Classification loss}
\todo {Write this out. What to do with the cost here?}

\subsection{Updating the belief state}
While the value function can learn certain dependencies between class occurrences, its job is made easier if the belief state already presents features that clearly present the relevant signals, such as co-occurence.
Toward this end, we employ an inference model to infer the marginals $P(C_i)$ given the observations.
It may also be useful to predict the marginals $P(L_i)$, as some formulations of our value function make use of this information.

We explore two alternatives for the inference model: a fully-connected Markov Random Field (MRF) and a locally weighted regression, as shown in Figure~\ref{fig:models}.

\begin{figure}[h!]
\centering
\subfloat[MRF]{\includegraphics[width=0.46\linewidth]{figures/inf_model_mrf.pdf}} \hfill
\subfloat[Local]{\includegraphics[width=0.46\linewidth]{figures/inf_model_loc.pdf}}
\caption{The models present a point in the middle of the decision process, when some classifiers have already been observed.}
\label{fig:models}
\end{figure}

\subsubsection{MRF inference}
Pros: (probably) more efficient at test time.

Cons: model too restrictive? not exact inference.

\subsubsection{Locally weighted regression}
Approach of \cite{Gao2011}.

Pros: imposing minimal model constraints (except Gaussian assumptions);

Cons: expensive at test time; data too sparse?

\subsection{Reward} \label{sec:reward}
Let's say that given deadline $T_d$ and some image, the policy had time to take $J$ actions.
The total expected reward of a policy $\pi$ is then defined as the sum of rewards
\begin{equation}
\mathbb{E}_\pi[\sum_{j=0}^J R(b_j)]
\end{equation}

The reward function does not necessarily have to be related to the evaluation of the system.
We explore two options for the reward function: the first is derived directly from the evaluation metric; the second seeks to minimize entropy of $P(\mathbf{C})$.

\subsubsection{Slope of performance vs. time}
\begin{figure}[htb]
  \center{\includegraphics[width=0.66\linewidth]{figures/apvst_expl.pdf}}
  \caption{\label{fig:rewards}A per-action greedy value function that corresponds to the maximization of our objective function is the area of the horizontal slice under the curve due to the action. The figure shows this analysis for the action highlighted in orange.}
\end{figure}

@@

Learning this accurately is the domain of Reinforcement Learning research, and is generally intractable for large-scale POMDPs.
We explore three different value functions.

@@

\subsubsection{Greedy expected-value}
We can compute the expected value of the reward function given a distribution over the classifier response.
If we define the value function by this expectation, we will have constructed a one-step greedy policy.

\note{Discuss submodularity defense of greedy policies here; costs are tricky though.}

@@

>>>>

% Definitions
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\def\subsectionautorefname{section}
\definecolor{light-gray}{gray}{0.5}
\newcommand{\aside}[1]{\textcolor{light-gray}{\emph{#1}}}
\newcommand{\todo}[1]{\textcolor{red}{\emph{#1}}}
\newcommand{\cut}[1]{\textcolor{light-gray}{#1}}
\newcommand{\comment}[1]{}

@@

We deal with a dataset of images $\mathcal{D}$, where each image $\mathcal{I}$ contains at least one, and often multiple, objects.
Each object is labeled with exactly one category label $k \in \{1, \dots, K\}$.

The multi-class, multi-label \emph{classification} problem asks whether $\mathcal{I}$ contains at least one object of class $k$.
The answer for a single label is given with a real-valued confidence by a function $\emph{classify}(\mathcal{I},k)$.
We write the ground truth for an image as $\mathbf{C}=\{C_1,\dots,C_K\}$, where $C_k \in \mathbb{B} = \{0,1\}$ is set to $1$ if an object of class $k$ is present.

@@

\cut{
We can make the classification problem more difficult by posing the \emph{counting} problem, which asks how many objects of class $k$ are present in $\mathcal{I}$, for each $k$.
This setting is not commonly evaluated; we mention it for its usefulness in later exposition.
}

The \emph{detection} problem is to output a list of bounding boxes (sub-images defined by four coordinates), each with a real-valued confidence that it encloses a single instance of an object of class $k$, for each $k$.
The answer for a single class is given by an algorithm $\emph{detect}(\mathcal{I},k)$, which outputs a list of sub-image bounding boxes $B$ and their associated confidences.

To highlight the hierarchical structure of these problems, we note that (1) the confidences may be given by $\operatorname{classify}(b,k)$ for each sub-image $b \in B$; (2) the correct answer to the detection problem also answers the counting and classification problems.

The following exposition deals with the classification problem, although we evaluate performance on both the classification and detection tasks.

@@

\begin{table}
\centering
\caption{Summary of our notation.}
\label{tab:notation}
\begin{tabular}{|l|l|}
	\hline
	$\mathcal{I}$	&	image \\
	$C_k$         & presence of class $k \in \{1,\dots,K\}$ \\ 
	$t$           & time into episode \\ 
	$T_s$, $T_d$  & start and deadline times \\ 
	$b_j$        	& belief state at step $j$ \\ 
	$\pi$         & policy function, $b \mapsto a \in \mathcal{A}$ \\
	$\mathcal{A}$ & set of actions $a_i \in \{\mathcal{L} \cup \mathcal{F}\}$\\ 
	$\mathcal{F}$ & set of featurization actions \\
	$\mathcal{L}$ & set of classification actions\\
	$o_i$					&	a real-valued observation upon executing $a_i \in \mathcal{L}$\\
	$\mathcal{O}$	& set of executed actions\\
	$\mathbf{o}$	& set of observations $\{o_k | a_k \in \mathcal{O}\}$\\
	$c(a)$				& cost of executing $a$, in units of $t$\\
	\hline
\end{tabular}\end{table}

@@

Our goal is a multi-class classification policy $\pi$ that takes an image $\mathcal{I}$ and outputs $\{\emph{classify}(1), \dots, \emph{classify}(K)\}$.
The policy repeatedly selects an action $a_i$ from a set of actions $\mathcal{A}$, executes it, potentially receives an observation $o_i$, and selects the next action.

A dynamic, or ``closed-loop,'' policy bases action selection on observations received from previous actions, exploiting the signal in inter-object and scene context for a maximally efficient path through the actions.
This is our goal, and what sets our formulation apart from multi-class systems that evaluate in a fixed order, such as simple cascades \cite{Viola2001} or \todo{add another one}.

The set of actions $\mathcal{A}$ can include classifiers and feature computations.
For simplicity of exposition, let $\mathcal{A}$ consist of $2K+2$ actions:
\begin{itemize}
	\item $F^\text{co}$, a complex spatial pyramid (SP) featurization of the image
	\item $K$ classifiers $L^\text{co}_i$: one-vs-all SVMs trained on the SP feature
	\item $F^\text{sc}$, a quick GIST featurization of the image
	\item $K$ classifiers $L^\text{sc}_i$: one-vs-all SVMs on the GIST feature
\end{itemize}

@@

The system is given two times: the setup time $T_s$ and deadline $T_d$.
Between these times, the system must strive for the goal of \emph{Anytime} performance and give the best possible answer at any point.

@@

To meet this goal, the evaluation metric of a policy is defined as the area under the performance vs. time curve between points $T_s$ and $T_d$ (see Figure~\ref{fig:evaluation}).
We evaluate policies by this more robust metric and not simply by the final performance at deadline time for the same reason that Average Precision is used instead of a fixed Precision vs. Recall point.

An action $a$ that consists of running a classifier $L_k$ returns a real-valued observation $o_k \sim P(O_k)$.
The state records the fact that $a$ has been taken by adding it to the initially empty set $\mathcal{O}$.
We refer to the current set of observations as $\mathbf{o} = \{o_k | L_k \in \mathcal{O}\}$.

We define the state $b$ of the decision process by the the distribution over class presence variables $P(\mathbf{C}) = P(C_1, \dots, C_K)$\footnote{We write $P(C_k)$ to mean $P(C_k=1)$.}, the time into episode $t$, and the set of executed actions $\mathcal{O}$ and corresponding observations $\mathbf{o}$.

A classification \emph{episode} takes an image $\mathcal{I}$ and proceeds from the initial belief state $b_0$ and action $a_0$ to the next pair $b_1$, and so on until $t$ exceeds $T_d$.
At that point, the policy is terminated and a new episode begins on a new image.

The policy's performance is determined by treating the final values of $P(\mathbf{C})$ as the set of confidence scores $\{\emph{classify}(1), \dots, \emph{classify}(K)\}$ for the multi-label classification evaluation described in \autoref{sec:problem_setting}.

For ease of reference, our notation is summarized in \autoref{tab:notation}.

\section{Selecting actions} \label{sec:value}
Optimal performance results from becoming maximally certain of the correct values $C_i$ as quickly as possible (given the setup time).
As our goal is to pick actions, we want to formulate a function $V(b,a)$ 	that assigns a value to a potential action, given the current state of the decision process.
We can then define the policy as simply
\begin{align}
\pi(b) = \argmax_{a \in \mathcal{A} \setminus \mathcal{O}} V(b,a)
\end{align}

Our system will be evaluated by some properties of the state $b$, and its performance is time-sensitive.
Accordingly, the value function should balance two terms: a function of the expected next belief state $b'$ after taking an action $a_i$, and the cost of $a_i$.
\begin{align}
V(b,a) = \mathbb{E}_{b' \sim P(b'|b,a)}[R(b')] - \lambda c(a|b)
\end{align}

We can be more explicit and split $b$ into its constitutents: (1) current set of observations $\mathbf{o}$; (2) posterior distribution $P(\mathbf{C}|\mathbf{o})$.
Accordingly, we have considerable flexibility in definining the ``reward'' function $R(b)$.

We consider a formulation of $R(b)$ based on information gain, as is common in problems of this type \todo{cite}.
The corresponding $R: P \mapsto \mathbb{R}$ is a function on the distribution over $\mathbf{C}$.
We note that as our goal is to reduce uncertainty of $P(\mathbf{C})$, it is natural to define $R$ as the negative entropy of the distribution given the observations.
\begin{align}
V(b,a)
&= \mathbb{E}_{o_i \sim P(O_i|\mathbf{C},\mathbf{o})}[R(P(\mathbf{C}|\mathbf{o} \cup o_i)] - \lambda c(a_i|\mathcal{O}) \\
&= \mathbb{E}_{o_i \sim P(O_i|\mathbf{C},\mathbf{o})}[-H(\mathbf{C}|\mathbf{o} \cup o_i)] - \lambda c(a_i|\mathcal{O})
\end{align}

Keeping in mind that $ H(\mathbf{C}|\mathcal{O}) = -\sum_{\mathbf{c}} P(\mathbf{c}|\mathcal{O}) \log P(\mathbf{c}|\mathcal{O}) $ is constant w.r.t $a_i$, it is easy to show that the policy consists of picking the action that maximizes information gain of $P(\mathbf{C})$, balanced with the cost of the action:
\begin{align}
\pi(b)
&= \argmax_{a_i \in \mathcal{A} \setminus \mathcal{O}} \, \mathbb{E}_{o_i \sim P(O_i|\mathbf{C},\mathbf{o})}[-H(\mathbf{C}|\mathbf{o} \cup o_i)] + H(\mathbf{C}|\mathbf{o}) - H(\mathbf{C}|\mathbf{o}) - \lambda c(a_i|\mathcal{O}) \\
&= \argmax_{a_i \in \mathcal{A} \setminus \mathcal{O}} \, I(\mathbf{C};O_i|\mathbf{o}) - H(\mathbf{C}|\mathbf{o}) - \lambda c(a_i|\mathcal{O}) \\
&= \argmax_{a_i \in \mathcal{A} \setminus \mathcal{O}} \, I(\mathbf{C};O_i|\mathbf{o}) - \lambda c(a_i|\mathcal{O})
\end{align}

\subsection{The model of the belief state}
\begin{figure}[h!]
\centering
\subfloat[MRF]{\includegraphics[width=0.46\linewidth]{figures/inf_model_mrf.pdf}}
\subfloat[Local]{\includegraphics[width=0.46\linewidth]{figures/inf_model_loc.pdf}}
\caption{We present a point in the middle of the decision process, when some classifiers have already been observed.}
\label{fig:model}
\end{figure}

Evaluation of our policy depends on being able to compute certain marginals for the value function.
Toward this end, we employ an inference model to infer the marginals $P(C_i)$ given the observations $\mathcal{O}$.
It may also be useful to predict the marginals $P(L_i)$, as some formulations of our value function make use of this information.

We employ a fully-connected Markov Random Field (MRF), as shown in Figure~\ref{fig:model}.
The parameters are trained on fully-observed data.
Exact inference is of course intractable in this model; we make use of Loopy Belief Propagation, which provides no general convergence guarantees but has been shown to work well empirically \todo{cite}.

\subsection{Submodularity} \label{sec:submodularity}
A natural question regarding our exposition so far is, why should we assume that a greedy policy would work well at all?
In fact, the problem as described exhibits interesting properties that here allow us to show that the greedy solution is near-optimal.

We first introduce the notion of submodularity.

\todo{write this out}

\subsection{Policy parametrization and more reward functions}
\todo{here introduce the view of learning parameters given the observed rewards}

Let's say that given deadline $T_d$ and some image, the policy had time to take $J$ actions.
The total expected reward of a policy $\pi$ is then defined as the sum of rewards
\begin{equation}
\mathbb{E}_\pi[\sum_{j=0}^J R(b_j)]
\end{equation}

The reward function does not necessarily have to be related to the evaluation of the system.

\subsubsection{Slope of performance vs. time}
\begin{figure}[htb]
  \centering
  \includegraphics[width=0.56\linewidth]{figures/apvst_expl.pdf}
  \caption{A per-action greedy value function that corresponds to the maximization of our objective function is the area of the horizontal slice under the curve due to the action. The figure shows this analysis for the action highlighted in orange.}
  \label{fig:rewards}
\end{figure}

@@

Learning this accurately is the domain of Reinforcement Learning research, and is generally intractable for large-scale POMDPs.

@@

@@

\section{Implementation Details and Evaluation}
We use FastInf \cite{Jaimovich2010} for the learning and inference in our MRF model.
